---
title: "Basic functionalities to deal with simulated mobile network data"
author: "Bogdan Oancea"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic functionalities to deal with simulated mobile network data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette contains a short introduction to *simutils* package. It describes its main purpose, presents some technical details of its implementation and provide examples on how to use this package. Some basic knowledge about the *simulator* tool is recommended. A detailed description of the data sets used by this package can be found in @epj, @rsr, @WPI.2, @WPI.3, @WPI.4, @arxiv and in @bmc_paper. To fully understand the structure of the data sets used by this package it is recommended to read the above mentioned papers.


# Introduction
This section contains a brief explanation about the intended use of the package. 

In the context of the European project ESSnet on Big Data, a mobile network data simulator has been developed in C++ @WPI.2, whose constant evolution is available as open source software in github: https://github.com/bogdanoancea/simulator. This simulator needs a set of input files in XML format and a geographical map in WKT format. These files contain data about:

* the simulation (time span information, mobility patterns, probabilities for individuals of having one or two mobile devices) 
* persons - how many individuals comprise the population, how fast they walk or move by car, how long they spend at home, at work, etc. 
* antennas - their position, emission power, path loss exponent, and other configuration parameters, and (iv) prior specifications for the computation of location probabilities.

Once the simulator is executed with all these parameters, the following outputs can be obtained as csv files: 

* the grid of analysis in which the geographical territory is divided 
* the ground truth about the persons at each time instant (their positions and their devices) 
* the parameters of each antenna 
* the network events arising from the connections of the devices to each antenna at each time instant
* the coverage area for each antenna 
* the signal strength (or the signal dominance) of each antenna in the center of each grid tile.

The structure of each csv file is specified in a corresponding dictionary in XML, whose contents, in turn, are specified also in a corresponding XSD file. These file structures allow the user to represent the metadata by using a standard format in XML as well as to conduct their validation with the dictionaries in XSD format. This validation is done using a Java program integrated in the *simutils* R package through the *rJava* package (@rj) allowing for compatibility with some features introduced in XML ver. 1.1 (features not yet in R packages dealing with XML processing).


Although the simulator has been implemented in C++, all the information related to it can be managed with the R language by using the *simutils* package. With this tool, it is possible to execute the simulator as well as to manage, specify and change their inputs and outputs. More, the outputs of the *simutils* package can be sent to *simviz*, another R package that we created to build different visualizations of the individual or aggregated data regarding the mobile devices and persons or have a visual representation of different aspects of the mobile network and its characteristics.


# Validation functions
The first step to run a simulation is to specify and validate the input files for the simulator software. We provide a set of XML input files for the simulator in the _extdata/input_files_ folder (simulation.xml, persons.xml, antennas.xml) together with the schema definition for each XML file found in the _extdata/input_files/schema_definition_ folder (simulation_dict.xsd, persons_dict.xsd, antennas_dict.xsd).

*simutils* package provides *validate_xml* function to validate an XML file against its corresponding XSD file.

```{r}
library(simutils)
rootPath <- file.path(system.file(package = "simutils"), 'extdata')  
	 
# Validate the simulation.xml file with its corresponding xsd file
xml_fn <- file.path(rootPath, "input_files", "simulation.xml")
xsd_fn <- file.path(rootPath, 
	                    "metadata/input_files/schema_definition", 
	                    "simulation_dict.xsd")
validate_xml(xsd_fn, xml_fn)

# Validate the persons.xml file with its corresponding xsd file
xml_fn <- file.path(rootPath, "input_files", "persons.xml")
xsd_fn <- file.path(rootPath, 
	                    "metadata/input_files/schema_definition", 
	                    "persons_dict.xsd")

	
# Validate the antennas.xml file with its corresponding xsd file
xml_fn <- file.path(rootPath, "input_files", "antennas.xml")
xsd_fn <- file.path(rootPath, 
	                    "metadata/input_files/schema_definition", 
	                    "antennas_dict.xsd")
validate_xml(xsd_fn, xml_fn)

```
# Updating simulation parameters

# Run a simulation

# Reading the simulation results and building a data structure suited for analysis

# Data Analysis

## Computing ground-truth totals

## Computing Origin Destination Matrices

# References
